Mendelian randomization for causal inference in Genome-wide association studies

### 2022-02-15 mr-review.Rmd
started writing applications section

```{r, dag2, fig.cap = "Scenario where the genetic variant G affects exposure X, which in turn affects Y. Unobserved confounder U affects both X and Y."}
d1 <- dagitty('dag {
bb="0,0,1,1"
G [pos="0,0"]
U [pos="0.75,10"]
X [exposure, pos="0.5,0"]
Y [outcome, pos="1.0,0"]
G -> X
U -> X
U -> Y
X -> Y
}'
)
plot(d1)

@@

To illustrate their approach, @burgess_sensitivity_2017 shared a case study in which they examined the causal effect of 
C-reactive protein (CRP) levels on coronary artery disease risk with four genetic 
variants in the CRP gene region and 17 other genetic variants that affect coronary 
artery disease risk. They use measured covariates to probe for SNP-covariate 
associations, and followed it up with scatter plots and Cochran's Q test on the causal 
estimates to inquire about whether the 
SNP instruments all identify the same causal parameter [@greco_m_detecting_2015;@small_sensitivity_2007;@higgins_measuring_2003]. 
Additionally, @burgess_sensitivity_2017 suggests using a funnel plot, like those in the 
meta-analysis literature, to visualize possible evidence of directional pleiotropy, where the average pleiotropic effects of the SNP instruments is nonzero [@sterne_recommendations_2011]. Additionally, Egger regression can be useful in this 
setting [@egger_bias_1997;@bowden_mendelian_2015].

@@

and $\tau$ is a vector of independent and identically distributed normal random errors with mean zero.

@@

>>>>

```{r, dag2, fig.cap = "Scenario where the genetic variant G affects exposure X, which in turn affects Y. Unobserved confounder U affects both X and Y."}
d1 <- dagitty('dag {
bb="0,0,1,1"
G [pos="0.0,0.0"]
U [pos="0.75,0.5"]
X [exposure, pos="0.5,0.0"]
Y [outcome, pos="1.0,0.0"]
G -> X
U -> X
U -> Y
X -> Y
}'
)
plot(d1)

@@

To illustrate their approach, @burgess_sensitivity_2017 shared a case study in which they examined the causal effect of 
C-reactive protein (CRP) levels on coronary artery disease risk with four genetic 
variants in the CRP gene region and 17 other genetic variants that affect coronary 
artery disease risk. They used measured covariates to probe for SNP-covariate 
associations, and followed it up with scatter plots and Cochran's Q test on the causal 
estimates to inquire about whether the 
SNP instruments all identify the same causal parameter [@greco_m_detecting_2015;@small_sensitivity_2007;@higgins_measuring_2003]. 
Additionally, @burgess_sensitivity_2017 suggested using a funnel plot, like those in the 
meta-analysis literature, to visualize possible evidence of directional pleiotropy, where the average pleiotropic effects of the SNP instruments is nonzero [@sterne_recommendations_2011]. Additionally, Egger regression can be useful in this 
setting [@egger_bias_1997;@bowden_mendelian_2015;@burgess_interpreting_2017;@bowden_framework_2017].

@@

and $\tau$ is a vector of independent and identically distributed normal random errors with mean zero. Note also that $\tau$ in Equation \@ref(eq:lm4) is independent of $\epsilon$ in Equation \@ref(eq:lm3).

@@

Birthweight is thought to impact long-term susceptibility to adulthood diseases. This scientific 
hypothesis, sometimes called the "fetal origins of adult diseases" motivated work by @zeng_causal_2019 in 
which the authors used MR to identify causal relationships between lower birthweight and greater risks of 
coronary artery disease, myocardial infarction, and type 2 diabetes mellitus [@barker_fetal_1990;@lucas_fetal_1999;@barker_fetal_2002]. Specifically, 
@zeng_causal_2019 used 
inverse variance-weighted MR to examine a total of 21 adult diseases and 38 other complex traits. Three causal associations met the Bonferroni multiple hypothesis testing threshold 
[@bonferroni1936teoria]. The authors analyzed data from 37 GWAS with up to 340,000 subjects with European 
ancestry [@zeng_causal_2019]. They estimated odds ratios of 1.34, 1.30, and 1.41 for coronary artery 
disease, myocardial infarction, and type 2 diabetes mellitus, respectively. Sensitivity analyses revealed 
the robustness of these findings to possible MR assumptions violations, including pleiotropy, maternal 
effects, and reverse causation [@zeng_causal_2019]. Curiously, none of the 38 other complex traits 
appeared to mediate the causal associations, perhaps due to limited sample sizes for these traits [@zeng_causal_2019]. @zeng_causal_2019 argue that further 
studies, including consideration of socioeconomic status and incorporation of larger sample sizes, may 
reveal mediating factors for these causal relationships and, thus, provide biological and public health insights.

@zeng_causal_2019-1

@zeng_birth_2019
